---
title: "Project 1"
output: pdf_document
date: "2023-10-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 
```{r}
Data <- read.csv("diamonds4.csv",header=T)
```

```{r}
library(tidyverse)
library(MASS)
library(faraway)
ggplot(Data,aes(x=carat,y=price)) + geom_point()+
  geom_smooth(method = "lm", se=FALSE)
```

The scatterplot comparing carat and price looks exponential rather than linear. This would indicate that a transformation is needed, likely a log transformation.

```{r}
result <- lm(price ~ carat,data=Data)
par(mfrow = c(2, 2))
plot(result)
```

```{r}
library(MASS)
MASS::boxcox(result)
```

Based on the diagnostic plots as well as the boxcox plots, it seems that both assumptions 1 and 2 have been violated since the errors do not have mean 0 or constant variance based on the residual plot. If both assumptions are violated, the likely solution is to log transform both the x and y variables. The box-cox plot agrees with this since the confidence interval does not include 1, indicating the y variable must be transformed.


```{r}
ystar<-log(Data$price)
Data<-data.frame(Data,ystar)
ggplot(Data,aes(x=carat,y=ystar)) + geom_point()+
  geom_smooth(method = "lm", se=FALSE)

xstar<-log(Data$carat)
Data<-data.frame(Data,xstar)
ggplot(Data,aes(x=xstar,y=ystar)) + geom_point()+
  geom_smooth(method = "lm", se=FALSE)
```
The first graph shows a scatter plot with just the y variable being log transformed. This does not remedy the exponential relationship, so the x is also log transformed, which ends up fixing the linear regression assumption. 

```{r}
result2 <- lm(ystar ~ xstar,data=Data)
par(mfrow = c(2, 2))
plot(result2)
```

After log transforming both the x and y variables, the diagnostic plots also pass the assumption, with a errors boht having a mean of aorund 0 and constant variance.


```{r}
summary(result2)
```
Based on the results of the model with both the response and the predictor being log transformed, xstar, log(carat), is very significant with a p value much lower than 0.05. This means that the log of carat is significant in predicting the log of the price of the diamond. 

The regression equation is log(y) = 8.521 + 1.944log(x). This means that for every 1% increase in the carat size, the price of the diamond is expected to increase by around 1.944%.


# Data Descriptions
Citation: Blue Nile. “The 4Cs of Diamonds,” n.d. https://www.bluenile.com/education/diamonds.
